---
title: "project"
author: "Ziyu Su"
date: "11/8/2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### load data
```{r echo=TRUE,eval=TRUE}
data=read.csv(file='diabetes.csv',header = TRUE)
```

### preprocessing
convert gender, frame to numeric values (gender: 0 for female, 1 for male; frame: 0 for small, 1 for medium, 2 for large):
```{r echo=TRUE,eval=TRUE}
a <- sub("female","0",data$gender)
b <- sub("male","1",a)
data$gender<-b
a=sub('small','0',data$frame)
b=sub('medium','1',a)
c=sub('large','2',b)
data$frame=c
data[data==""]=NA
```
remove id,location,bp.2s,bp2d; remove rows with NA value
```{r echo=TRUE,eval=TRUE}
data$id=NULL
data$location=NULL
data$bp.2d=NULL
data$bp.2s=NULL

data=na.omit(data)
```
366 data points remain

convert height and weight into h/w, convert waist and hip into ws/hp
```{r echo=TRUE,eval=TRUE}
data$h.w=data$height/data$weight
data$ws.hp=data$waist/data$hip
data$height=NULL
data$weight=NULL
data$waist=NULL
data$hip=NULL
```

create label (0 for healthy, 1 for diabetes):
```{r echo=TRUE,eval=TRUE}
label=data$glyhb
label[label<6.5]=0
label[label>=6.5]=1
```

### data exploration

```{r echo=TRUE,eval=TRUE}
plot(data$glyhb,data$stab.glu)
plot(data$glyhb,data$ws.hp)
```

```{r echo=TRUE,eval=TRUE}
hist(label)
hist(data$glyhb)
```

### t-SNE visualization

```{r echo=TRUE,eval=TRUE}
library(Rtsne)
## Curating the database for analysis with both t‐SNE and PCA
data0=data
data0$glyhb=NULL
data0$label=as.factor(label)
## for plotting
colors = rainbow(length(unique(data0$label)))
names(colors) = unique(data0$label)
## Executing the algorithm on curated data
tsne=Rtsne(data0[,-1], dims = 2, perplexity=30, verbose=TRUE, max_iter = 500)
exeTimeTsne=system.time(Rtsne(data0[,-1], dims = 2, perplexity=30, verbose=TRUE, max_iter = 500))

## Plotting
plot(tsne$Y, t='n', main="tsne")
text(tsne$Y, labels=data0$label, col=colors[data0$label])
```


### variable selection
"glyhb", which means Glycosolated Hemoglobin, is popularly used indicator of diabetes. Thus, we want to build a model to fit the "glyhb" value. We used step-wise regression to select variables to be used.
```{r echo=TRUE,eval=TRUE}
step(lm(glyhb~1, data=data), direction="both", scope=~chol+stab.glu+hdl+ratio+age+gender+frame+bp.1s+bp.1d+time.ppn+h.w+ws.hp)
```


| Variables    | (Intercept) |  stab.glu |    chol   |    age    |   ratio   |  time.ppn |
| ------------ | ----------- | --------- | --------- | --------- | --------- | --------- |
| Coefficients | 0.2767223   | 0.0276492 | 0.0043767 | 0.0150447 | 0.1239409 | 0.0005546 |


$\hat{y}=0.2767223+0.0276492x_1+0.0043767x_2+0.0150447x_3+0.1239409x_4+0.0005546x_5$

therefore, we select to use stab.glu (Stabilized Glucose), chol (Total Cholesterol), age, ratio(Cholesterol/HDL Ratio), time.ppn (Postprandial Time when Labs were Drawn)

### model fitting

```{r echo=TRUE,eval=TRUE}
set.seed(100)
index <-sort(sample(nrow(data), nrow(data)*.8))
train_data <-data[index,]
train_label=label[index]
test_data <-data[-index,]
test_label=label[-index]
```

```{r echo=TRUE,eval=TRUE}
data0=data
data0$hdl=NULL
data0$age=NULL
data0$gender=NULL
data0$frame=NULL
data0$bp.1s=NULL
data0$bp.1d=NULL
data0$h.w=NULL
data0$ws.hp=NULL
train_data0 <-data0[index,]
train_label0=label[index]
test_data0 <-data0[-index,]
test_label0=label[-index]

fit0=lm(glyhb~., data = train_data0)
summary(fit0)
```

```{r eval=TRUE,echo=TRUE}
a=test_data0
test_y0=a$glyhb
a$glyhb=NULL
test_x0=a
```

```{r eval=TRUE,echo=TRUE}
test_pre0=predict(fit0,test_x0,interval = 'prediction')


test_pre0[test_pre0<6.5]=0
test_pre0[test_pre0>=6.5]=1
```

```{r eval=TRUE,echo=TRUE}
b=data.frame(test_pre0)
a=sum(abs(b$fit-test_label0))
table1=table(b$fit,test_label0)
table1
acc=1-a/nrow(test_data0)
acc
plot(fit0)
```

```{r echo=TRUE,eval=TRUE}
fit1=lm(glyhb~.,data = train_data)
summary(fit1)
```

```{r eval=TRUE,echo=TRUE}
plot(fit1)
```

```{r eval=TRUE,echo=TRUE}
a=test_data
test_y=a$glyhb
a$glyhb=NULL
test_x=a
```

```{r eval=TRUE,echo=TRUE}
test_pre=predict(fit1,test_x,interval = 'prediction')
test_pre
```
```{r eval=TRUE,echo=TRUE}
test_pre[test_pre<6.5]=0
test_pre[test_pre>=6.5]=1
```

```{r eval=TRUE,echo=TRUE}
b=data.frame(test_pre)
a=sum(abs(b$fit-test_label))
table1=table(b$fit,test_label)
table1
acc=1-a/nrow(test_data)
acc
```
|Predict\Label|  0  |  1  |
| ----------- | --- | --- |
|      0      | 60  |  4  |
| ----------- | --- | --- |
|      1      |  2  |  8  |
### SVM
Prepare data2 for SVM. Add label into it and remove the glyhb.
```{r eval=TRUE,echo=TRUE}
data2=data # data for svm
data2$label=factor(label)
data2$glyhb=NULL
set.seed(100)
index <-sort(sample(nrow(data2), nrow(data2)*.8))
train_data2 <-data2[index,]
test_data2 <-data2[-index,]
```


```{r eval=TRUE,echo=TRUE}
library(e1071)
svm1=svm(label~.,data=train_data2)
table2=table(predict(svm1,test_data2[,-13]),test_data2[,13])
table2
acc=sum(diag(table2))/nrow(test_data2)
acc
```
























